Methods with preconditioning with weight decay regularization

### 2023-03-21 paper/main.tex
add two theorems

@@

>>>>

\newtheorem{theorem}{Theorem}

@@

\section{Theory}
To proof this algorigthm we use simplified form of AdamW algorithm
\begin{algorithm}[H]
\caption{AdamW}\label{alg:AdamW}

    \begin{algorithmic}
    \Require{$r, \varepsilon, f$}    
    \While {$w$ not converged}
        \State $t = t+1$
        \State $D_t = \text{diag}(|\nabla f(w_t)|_i)$
        \State $w_t = w_{t-1} - \eta \cdot \nabla f(w_t) D_t^{-1}  -\lambda \nabla r(w_t)$ 
    \EndWhile
\end{algorithmic}
\end{algorithm}

\begin{theorem}

Suppose the Assumption 1, 2, 5 and let $\varepsilon > 0$ and let the step-size satisfy
\begin{equation*}
    \eta < \frac{2 \alpha}{L + l \cdot \alpha} 
\end{equation*}
Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem (link here to problem 1) can be bounded by
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}

\end{theorem}
\begin{proof}

Let's write first assumption for step t and $t+1$:

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \nabla f(w_t), w_{t+1} - w_t \rangle + \frac{L}{2}||w_{t+1} - w_t ||^2
\end{equation*}
Okay, by definition for our algorithm we have:

\begin{equation*}
w_{t+1} - w_t = -\eta D_t^{-1} \nabla f(w_t) - \eta \nabla r(w_t)    
\end{equation*}

and

\begin{equation*}
\nabla f(w_t) = \frac{1}{\eta} D^t(w_t - w_{t+1}) - D^t \nabla r(w_t)
\end{equation*}

Okay, now let's replace $\nabla f(w_t)$ and $I \leq \frac{D_t}{\alpha}$
\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \langle \frac{1}{\eta}D_t(w_t - w_{t+1}) - D_t\nabla r(w_t), w_{t+1} - w_t \rangle + \frac{L}{2 \alpha} ||w_{t+1} - w_t||_{D_t}^2
\end{equation*}

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left(\frac{L}{2 \alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 - \langle D_t \nabla r(w_t), w_{t+1} - w_t \rangle
\end{equation*}

Lets define new variable $\tilde{r} : \nabla \tilde{r} = D_t \nabla r(w_t)$. Then rewrite step using the variable and 5-th assumption.
\begin{equation*}
    \tilde{r}(w_{t+1}) \leq \tilde{r}(w_t) + \langle \tilde{r}(w_t), w_{t+1} - w_t \rangle + \frac{l}{2} (w_{t+1} - w_t)^T D_t (w_{t+1} - w_t)
\end{equation*}

\begin{equation*}
    f(w_{t+1}) \leq f(w_t) + \left( \frac{L}{2\alpha} - \frac{1}{\eta} \right) ||w_{t+1} - w_t||_{D_t}^2 + \tilde{r}(w_t) - \tilde{r}(w_{t+1}) + \frac{l}{2}||w_{t+1}-w_t||_{D_t}^2
\end{equation*}

$\tilde{F}(w) = f(w) + \tilde{r}(w)$, $F(w) = f(w) + r(w)$, ($\tilde{L}=L + l \alpha$), we get:

\begin{equation*}
    \tilde{F}(w_{t+1}) \leq \tilde{F}(w_t) + \left( \frac{\tilde{L}}{2\alpha} - \frac{1}{\eta}  \right) ||w_{t+1} - w_t||_{D_t}^2
\end{equation*}

\begin{equation*}
    \left(\frac{1}{\eta} - \frac{\tilde{L}}{2\alpha}   \right) ||w_{t+1} - w_t||_{D_t}^2 \leq \tilde{F}(w_t) - \tilde{F}(w_{t+1})
\end{equation*}

\begin{equation*}
    \frac{\eta^2  (T+1)}{\Gamma}\left(\frac{1}{\eta} - \frac{\tilde{L}}{2\alpha}   \right)\cdot\min_{k = 0, T} ||\nabla f(w_t) + \nabla \tilde{r}(w_t)||^2 \leq \frac{\eta^2}{\Gamma}\left(\frac{1}{\eta} - \frac{\tilde{L}}{2\alpha}   \right)\cdot\sum\limits_{t = 0}^T ||\nabla f(w_t) + \nabla \tilde{r}(w_t)||^2 \leq \tilde{F}(w_0) - \tilde{F}(w_*)
\end{equation*}

\begin{equation*}
    \min_{t = 0, T} ||\nabla f(w_t) + \nabla \tilde{r}(w_t)||^2 \leq \frac{(\tilde{F}(w_0) - \tilde{F}(w_*))\Gamma}{(\frac{1}{\eta} - \frac{\tilde{L}}{2\alpha}) \eta^2 (T+1)} = \varepsilon
\end{equation*}

\begin{equation*}
    T + 1 \geq \frac{\Delta_0 \Gamma}{(\frac{1}{\eta} - \frac{\tilde{L}}{2\alpha}) \eta^2 \varepsilon}
\end{equation*}

Then:
\begin{equation*}
      T = \mathcal{O}\left( \frac{2\Delta_0 \Gamma \alpha } {(2\alpha - \tilde{L}\eta) \eta \varepsilon} \right)
\end{equation*}
\end{proof}

\begin{theorem}
    Suppose the Assumption 1, 2, 4, 5 and let $\varepsilon > 0$ and let the step-size satisfy
    \begin{equation*}
        \eta \leq \frac{2 \alpha}{\tilde{L}}
    \end{equation*}
    Then, the number of iterations performed by AdamW algorithm, starting from an initial point $w_0 \in \mathbb{R}^d$ with $\Delta_0 = \tilde{F}(w_0) - \tilde{F}^*$, required to obtain and $\varepsilon$-approximate solution of the convex problem (link here to problem 1) can be bounded by
    \begin{equation*}
        T =  \mathcal{O}\left( \frac{\ln \frac{\Delta_0}{\epsilon}}{2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha})} \right)
    \end{equation*}
\end{theorem}

\begin{proof}
    Assume 
    \begin{equation}
    \nabla \tilde{F} = \nabla f + \nabla \tilde{r}    
    \end{equation}

    \begin{equation}
    L + ||D_t||l = \tilde{L}        
    \end{equation}
    \begin{equation*}
    w_{t+1} - w_t = -\eta D_t^{-1} \nabla r(w_t) - \eta \nabla r(w_t) = -\eta D_t^{-1} (\nabla f + \nabla \tilde{r})(w_t) = -\eta D_t^{-1} \nabla \tilde{F}(w_t)    
    \end{equation*}
    Then we write $\tilde{L}$-smoothness for $\tilde{F}$ 
    \begin{equation*}
        \tilde{F}(w_{t+1}) - \tilde{F}(w_t) \leq  \langle \nabla \tilde{F}(w_t), w_{t+1} - w_t \rangle + \frac{\tilde{L}}{2} ||w_{t+1} - w_t||^2
    \end{equation*}
    \begin{equation*}
        \tilde{F}(w_{t+1}) - \tilde{F}(w_t) \leq - \langle \frac{1}{\eta} D_t(w_{t+1} - w_t), w_{t+1} - w_t \rangle + \frac{\tilde{L}}{2} ||w_{t+1} - w_t||^2 = (\frac{\tilde{L}}{2 \alpha} - \frac{1}{\eta}) ||w_{t+1} - w_t||^2_{D_t} 
    \end{equation*}
    \begin{equation*}
        = (\frac{\tilde{L}}{2 \alpha} - \frac{1}{\eta}) ||w_{t+1} - w_t||^2_{D_t} = (\frac{\tilde{L}}{2 \alpha} - \frac{1}{\eta}) ||-\eta D_t^{-1} \nabla \tilde{F}(w_t)||_{D_t}^2  \leq (\frac{\tilde{L}}{2 \alpha} - \frac{1}{\eta}) \eta^2 ||\nabla \tilde{F}(w_t)||^2_{D_t^{-1}}
    \end{equation*}
    Then we use PL-condition for the function $\tilde{F}$:
    \begin{equation*}
        ||\nabla \tilde{F}(w_t)||_{D_t^{-1}}^2 \geq 2 \mu (\tilde{F}(w_t) - \tilde{F}^*)
    \end{equation*}

    \begin{equation*}
    \tilde{F}(w_{t}) -  F^* \ge \tilde{F}(w_{t+1}) - \tilde{F}^* + (\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha}) \eta^2 2 \mu (\tilde{F}(w_t) - \tilde{F}^*) = \left( 1 +  2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha}) \right) (\tilde{F}(w_{t+1}) - \tilde{F}^*)        
    \end{equation*}

    \begin{equation*}
    \epsilon \ge \Delta_0 \left( 1 +  2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha}) \right)^{-T} \ge (\tilde{F}(w_{T}) - \tilde{F}^*)        
    \end{equation*}

    \begin{equation*}
    T = \frac{\ln \frac{\Delta_0}{\epsilon}}{\ln(1 + 2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha}))} \approx \frac{\ln \frac{\Delta_0}{\epsilon}}{2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha})}        
    \end{equation*}
    Then:
    \begin{equation*}
    T =  \mathcal{O}\left( \frac{\ln \frac{\Delta_0}{\epsilon}}{2 \mu \eta^2(\frac{1}{\eta} - \frac{\tilde{L}}{2 \alpha})} \right)
    \end{equation*}
\end{proof}
